---
title: "Assignment 3"
author: "Bryanna Schaffer"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing necessary packages:
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(pastecs)
```


Importing the dataset (TextMessages):

```{r}
TextMessages <- read_csv("TextMessages.csv")
names(TextMessages) # shows variable names
head(TextMessages) # shows first 6 rows
```

Renaming Group variable for easier read:

```{r}
TextMessages <- TextMessages %>%
  mutate(Group = case_when(
    Group == 1 ~ "Group 1", 
    Group == 2 ~ "Group 2"
  )
  )

head(TextMessages$Group)
```

Converting data from wide to long format:

```{r}
TextMessages_long <- TextMessages %>%
  pivot_longer(
    cols = c(Baseline, Six_months),
    names_to = "Time",
    values_to = "Messages"
  )
names(TextMessages_long)
head(TextMessages_long)
```

Visualization #1: Stratified boxplots of text messages by Group and Time

```{r}
ggplot(TextMessages_long, aes(x = Time, y = Messages, fill = Group)) + geom_boxplot() + 
  facet_wrap(~ Group) +
  labs(
    title = "Text Messages by Group and Time",
    x = "Time (Baseline or Six Months)",
    y = "Number of Text Messages"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Summary Statistics: Summary statistics of Text Messages by Group and Time. I also performed summary statistics of Text Messages by Group and Text Messages by Time.

```{r}

by(TextMessages_long$Messages, TextMessages_long$Group, stat.desc)

by(TextMessages_long$Messages, TextMessages_long$Time, stat.desc)

summary <- by(TextMessages_long$Messages, 
   list(Group = TextMessages_long$Group, Time = TextMessages_long$Time), 
   stat.desc)

table(summary)

```
