---
title: "Assignment 3"
author: "Bryanna Schaffer and Ibrahim Elbasheer"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing necessary packages:
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(pastecs)
```


Importing the dataset (TextMessages):

```{r}
TextMessages <- read_csv("TextMessages.csv")
names(TextMessages) # shows variable names
head(TextMessages) # shows first 6 rows
```

Renaming Group variable for easier read:

```{r}
TextMessages <- TextMessages %>%
  mutate(Group = case_when(
    Group == 1 ~ "Group 1", 
    Group == 2 ~ "Group 2"
  )
  )

head(TextMessages$Group)
```

Converting data from wide to long format: This allows for Time to be represented in two categories: baseline and six_months. It takes those values and assigns it to that specific time period. Now, this long data set has the variables Group, Participant, Time, and Messages.

```{r}
TextMessages_long <- TextMessages %>%
  pivot_longer(
    cols = c(Baseline, Six_months),
    names_to = "Time",
    values_to = "Messages"
  )
names(TextMessages_long)
head(TextMessages_long)
```

Visualization #1 - Bryanna: Stratified boxplots of text messages by Group and Time

Below, we used the long format dataset to show text messages by group and time. To do so, we filled and wrapped our plot with the Group variable (Group 1 and Group 2). We see a similar trend of text messages over time for each group. There appears to be a decrease in the median number of text messages from baseline to six months after the baseline. However, it appears that Group 2 has a greater median number of text messages overall, meaning at the baseline and six months after. There are multiple outliers at the six month post baseline for group 1, while for group 2 there appears to only be two outliers. 

```{r}
ggplot(TextMessages_long, aes(x = Time, y = Messages, fill = Group)) + geom_boxplot() + 
  facet_wrap(~ Group) +
  labs(
    title = "Text Messages by Group and Time",
    x = "Time (Baseline or Six Months)",
    y = "Number of Text Messages"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Visualization #2 - Ibrahim: Stratified bar plot of text messages by Group and Time

The bar plot shows difference in means between  baseline and six months after the baseline period for each group. According to the plot, in this sample, the average number of text messagesis significantly higher in the baseline than six months after for both groups (Group 1 and Group 2).
```{r}
barplot_d_by_Group_and_Time<-ggplot(TextMessages_long,aes(x = Time,
                                                  y = Messages, fill = Group))
barplot_d_by_Group_and_Time+ 
  stat_summary(fun.data = mean_cl_normal,geom = "bar", colour = "blue" ,
               fill="lightblue") +
  labs(title="Text Messages by Group and Time Period",
       x="Average Number of Text Messages",
       y="Time") +
  facet_wrap(.~Group, nrow = 1)
```


Summary Statistics - Bryanna: Summary statistics of Text Messages by Group and Time. I also performed summary statistics of Text Messages by Group and Text Messages by Time.

Based on the summary statistics of text messages by both group and time, we see the greatest mean text messages to be 65.6 in the baseline measurement of group 2. We see the minimum value of text messages in group 1 six months after the baseline is 9, which is seen as an outlier in the boxplot above. Group 1 six months after the baseline has the greatest amount of variation of approximately 266.71  with a range of 69, while group 2 six months after the baseline has the least amount of variation of 88.56 with a range of 33. This also follows the greatest standard deviation of 16.33 in group 1 six months after baseline, while for group 2 six months after baseline the standard deviation is 9.41.

```{r}

by(TextMessages_long$Messages, TextMessages_long$Group, stat.desc)

by(TextMessages_long$Messages, TextMessages_long$Time, stat.desc)

summary <- by(TextMessages_long$Messages, 
   list(Group = TextMessages_long$Group, Time = TextMessages_long$Time), 
   stat.desc)

summary

```
